---
title: "TRGN510-Final-Project"
author: "Quinta Wang"
date: "2022-12-05"
output: html_document
---

Install Deseq2
```{r}
library(BiocManager)
BiocManager::install("DESeq2")
```

```{r}
library(DESeq2)
library(ggplot2)
```
Load all data including the gene id, unstranded data and sample name and condition.
```{r}
data <- read.delim("~/Desktop/backup/merge_file.txt", header = TRUE, sep = ",")
samplename  <- read.delim("~/Desktop/TRGN510FinalProjectTSV - Sheet1.tsv", header = TRUE, sep = "\t")
```

```{r}
samplename$Condition <- as.factor(samplename$Condition)
```

```{r}
row.names(data) <- data$gene_id
data <- data[-c(1)]
```


```{r}
DES_data <- DESeqDataSetFromMatrix(countData = data,
                                   colData = samplename,
                                   design = ~ Condition)
```


```{r}
DES_data <- DES_data[rowSums(counts(DES_data) > 10),]
```


```{r}
DES_data <- factor(DES_data$Condition, levels = c("Young","Senior"))
```

```{r}
condition <- factor(c("SN","SN","SN","SN"))
```

```{r}
dds <- DESeq(DES_dataset)
res <- results(dds)
res
```
# Log fold change shrinkage for visualization and ranking
```{r}
res <- results(dds, name="Condition_Young_vs_Senior")
res <- results(dds, contrast=c("Condition","Young","Senior"))
```


```{r}
resultsNames(dds)
```

```{r}
BiocManager::install("apeglm")
library(apeglm)
```

```{r}
resLFC <- lfcShrink(dds, coef="Condition_Young_vs_Senior", type="apeglm")
resLFC
```
# p-values and adjusted p-values
```{r}
resOrdered <- res[order(res$pvalue),]
```

```{r}
summary(res)
```
```{r}
sum(res$padj < 0.1, na.rm = TRUE)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r}
sum(res05$padj < 0.05, na.rm=TRUE)

```

# Exploring and exporting results
##MA-plot
```{r}
plotMA(resLFC, ylim=c(-2,2))
```
# Plot counts
```{r}
plotCounts(dds, gene=which.min(res$padj), intgroup="Condition")
```
```{r}
d <- plotCounts(dds, gene=which.min(res$padj), intgroup="Condition", 
                returnData=TRUE)
ggplot(d, aes(x=Condition, y=count)) + 
  geom_point(position=position_jitter(w=0.1,h=0)) + 
  scale_y_log10(breaks=c(25,100,400))
```

# Exporting results to CSV files
```{r}
write.csv(as.data.frame(resOrdered), 
          file="Condition_age_results.csv")
```

```{r}
resSig <- subset(resOrdered, padj < 0.1)
resSig
```
```{r}
vsd <- vst(dds, blind=FALSE)
rld <- rlog(dds, blind=FALSE)
head(assay(vsd), 3)
```

```{r}
BiocManager::install("vsn")
```

# Effects of transformations on the variance
```{r}
ntd <- normTransform(dds)
library("vsn")
meanSdPlot(assay(ntd))
```

```{r}
meanSdPlot(assay(vsd))
```
```{r}
meanSdPlot(assay(rld))
```
```{r}
BiocManager::install("pheatmap")
```

# Heatmap of the count matrix
# Heatmap with normal transform
```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("Sample.Name", "Condition")])
pheatmap(assay(ntd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)

```
# Heatmap with variance stablizing transformation
```{r}
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)

```

# Heatmap with regularized log transformation
```{r}
pheatmap(assay(rld)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```
# Heatmap of the sample-to-sample distances
```{r}
sampleDists <- dist(t(assay(vsd)))
DistMatrix <- as.matrix(sampleDists)
pheatmap(DistMatrix)
```
# Principal component plot of the samples
```{r}
plotPCA(vsd, intgroup=c("Condition"))
```

